Circumnuclear Media and Accretion Rates of Quiescent Supermassive Black Holes

### 2015-03-09 ms.tex
conclusion

Apologies: 
\begin{itemize}
\item{we assume that average star formation histories correspond to those experienced in the nuclear region.  }

\item{ Do nuclear star clusters in other galaxies look like scaled
    up/down versions of the one in our own galaxy? If so this would
    imply a radial gradient in the average of the stellar population
    which would imply a heating which would be dependent on radius.  }

\end{itemize}

However, \citet{McCourt+12} argue that even if cooling is exactly
balanced by heating everywhere, there is still a local thermal
instability that operates to turn hot gas into cold clouds.  However,
this local instability only causes a significant local drop-out of
mass from the hot medium (to low-T clouds), if the cooling timescale,
\begin{align}
\frac{\tcool}{\tff} &\equiv \frac{3\rho k T/2 \mu \mp}{\dot{q}_{\rm cool}} \approx \frac{9}{5} \left. \frac{q \vw^2/2}{\dot{q}_{\rm cool}}\right|_{\rs}
\end{align} 
is locally shorter than the approximately 10 times the free-fall
timescale $t_{\rm ff}$, where the last equality is evaluated at the
stagnation radius and makes use of equations~(\ref{eq:Tanalytic}) and
(\ref{eq:rhors2}).  We thus see that the condition $\tilde{v}_{w} \gtrsim v_{\rm
  crit}$ is also the condition for local instability, although the
prefactor may be slightly different.

If cooling becomes important in the flow ($\tilde{v}_{w} < v_{\rm CI}$), it may lead to thermal instabilities.  There are two kinds of ``instabilities", global and local.  The global instability is simply that, if cooling becomes very rapid, the gas loses thermal pressure and starts to inflow faster, i.e. the entire CNM accretes.  In
practice this should could manifest as a large increase in the stagnation radius, due an effective {\it decrease} in the effective heating rate.  This global ``instability" is not a true instability insofar as the gas does not necessarily clump locally into low-T clouds.

We have calculated steady-state models for circumnuclear medium of
early type galaxies, assuming that gas is supplied entirely by stellar
wind mass loss and heated by shocked stellar winds, supernovae and
black hole feedback. We may use these profiles to calculate an $\Mdot$
onto each of sample galaxy's SMBH, and get the scaling of $\Mdot$ with
$\Mbh$. We may then compare our results with observed trends of
nuclear x-ray luminosity. Our main conclusions may be summarized as
follows.

  \begin{enumerate}
  %Younger population=>higher accretion rate; generally higher eta and
  %lower vw except for very young stellar populations.
  \item We find that $\eddr \propto \Mbh^{\alpha}$ for
    $\Mbh\gsim4\times10^6$ and $\Mbh\lsim 2\times 10^8$, where the
    value of $\alpha$ with $\alpha\simeq$ for galaxies following the
    average star formation histories measured in cosmological
    simulations \citep{MosterNaab+:2013a}. This is in tension
    observations of nuclear X-ray luminosity which suggest that
    $\eddr$ increases towards smaller $\Mbh$
    \citep{Miller+15}. However, it is important to note that the
    inflow rate on large scale (which we calculate) may differ from
    the actual accretion rate as outflows could occur on small
    scales. The outflow rate would depend on the angular momentum of
    the infalling gas and thus the angular momentum of stars in the
    nucleus. The in turn would vary systematically with $\Mbh$.
  \item We expect within scatter in nuclear x-ray luminosities at
    fixed $\Mbh$ from (1) stochastic variation in star formation
    history from galay to galaxy (2) Variations in small scale outflow
    rates due to variations between galactic nuclear angular momenta.
    These variations alone could explain much of the 2-4 orders of
    magnitude of scatter of nuclear x-ray luminosity at fixed $\Mbh$ 
    in the samples of \citet{Miller+15} and \citet{Pellegrini:2010a}.

    It is also possible that many of the galaxies in the
    \citet{Miller+15} are accreting in a non-steady fashion and this
    scatter arises from cyclic variation in accretion rate. In fact it
    that a population of cyclically accreting with $\Mbh\gsim 10^{8}
    \Msun$ coexists with a population of smaller, steadily accreting
    black holes (e.g. the upper limits in \citet{Miller+15}.  This
    calls into question the use of simple extrapolations of power-law
    fits to the $L_{\rm X}(M_{\bullet}$) relationship to low $M_{\bullet}$
    to constrain the occupation fraction of black holes in low mass
    galaxies (\citealt{Miller+15}).
  \item Low mass AGN have growth times comparable to the Hubble
    time. We have show that this would require accretion rates not
    consistent with steady state thermally stable accretion of stellar
    wind material.
  \item We compare the $\Mdot$ for each of our solution to the one
    calculated from the Bondi formula. We find that they generally to
    a factor of a few. In our model the accretion rate is determined
    from the gas properties at the stagnation radius, $\rs$, which
    agrees with the Bondi radius, $\rb$, within a factor of a
    few. Thus, use of the Bondi formula is a good approximation.
  \end{enumerate}

>>>>

% Apologies: 
% \begin{itemize}
% \item{we assume that average star formation histories correspond to those experienced in the nuclear region.  }

% \item{ Do nuclear star clusters in other galaxies look like scaled
%     up/down versions of the one in our own galaxy? If so this would
%     imply a radial gradient in the average of the stellar population
%     which would imply a heating which would be dependent on radius.  }

% \end{itemize}

% However, \citet{McCourt+12} argue that even if cooling is exactly
% balanced by heating everywhere, there is still a local thermal
% instability that operates to turn hot gas into cold clouds.  However,
% this local instability only causes a significant local drop-out of
% mass from the hot medium (to low-T clouds), if the cooling timescale,
% \begin{align}
% \frac{\tcool}{\tff} &\equiv \frac{3\rho k T/2 \mu \mp}{\dot{q}_{\rm cool}} \approx \frac{9}{5} \left. \frac{q \vw^2/2}{\dot{q}_{\rm cool}}\right|_{\rs}
% \end{align} 
% is locally shorter than the approximately 10 times the free-fall
% timescale $t_{\rm ff}$, where the last equality is evaluated at the
% stagnation radius and makes use of equations~(\ref{eq:Tanalytic}) and
% (\ref{eq:rhors2}).  We thus see that the condition $\tilde{v}_{w} \gtrsim v_{\rm
%   crit}$ is also the condition for local instability, although the
% prefactor may be slightly different.

% If cooling becomes important in the flow ($\tilde{v}_{w} < v_{\rm CI}$), it may lead to thermal instabilities.  There are two kinds of ``instabilities", global and local.  The global instability is simply that, if cooling becomes very rapid, the gas loses thermal pressure and starts to inflow faster, i.e. the entire CNM accretes.  In
% practice this should could manifest as a large increase in the stagnation radius, due an effective {\it decrease} in the effective heating rate.  This global ``instability" is not a true instability insofar as the gas does not necessarily clump locally into low-T clouds.

We have calculated steady-state models for circumnuclear medium of
quiescent galaxies, assuming that gas is supplied entirely by stellar
wind mass loss and heated by shocked stellar winds, supernovae and
black hole feedback. We compute profile for a range of different black
hole masses, heating rates, and stellar density profiles and use these
profiles to calculate an $\Mdot$ onto each galaxy's SMBH, and get the
scaling of $\Mdot$ with $\Mbh$. We may then compare our results with
observed trends of Our main conclusions may be summarized as follows.

  \begin{enumerate}
  %Younger population=>higher accretion rate; generally higher eta and
  %lower vw except for very young stellar populations.
  \item We find that $\eddr \propto \Mbh^{}$ for
    $\Mbh\gsim4\times10^6$ and $\Mbh\lsim 2\times 10^8$, for galaxies
    following the average star formation histories measured in
    cosmological simulations \citep{MosterNaab+:2013a}. This is in
    tension observations of nuclear X-ray luminosity which suggest
    that $\eddr$ increases towards smaller $\Mbh$
    \citep{Miller+15}. However, it is important to note that the
    inflow rate on large scale (which we calculate) may differ from
    the actual accretion rate as outflows could occur on small
    scales. The outflow rate would depend on the angular momentum of
    the infalling gas and thus the angular momentum of stars in the
    nucleus. The in turn would vary systematically with $\Mbh$.
  \item We expect within scatter in nuclear x-ray luminosities at
    fixed $\Mbh$ from (1) stochastic variation in star formation
    history from galay to galaxy (2) Variations in small scale outflow
    rates due to variations between galactic nuclear angular momenta.
    These variations alone could explain much of the 2-4 orders of
    magnitude of scatter of nuclear x-ray luminosity at fixed $\Mbh$ 
    in the samples of \citet{Miller+15} and \citet{Pellegrini:2010a}.

    It is also possible that many of the galaxies in the
    \citet{Miller+15} are accreting in a non-steady fashion and this
    scatter arises from cyclic variation in accretion rate. In fact it
    that a population of cyclically accreting with $\Mbh\gsim 10^{8}
    \Msun$ coexists with a population of smaller, steadily accreting
    black holes (e.g. the upper limits in \citet{Miller+15}).  This
    calls into question the use of simple extrapolations of power-law
    fits to the $L_{\rm X}(M_{\bullet}$) relationship to low $M_{\bullet}$
    to constrain the occupation fraction of black holes in low mass
    galaxies (\citealt{Miller+15}).
  \item Low mass AGN have growth times comparable to the Hubble time
    \citet{Heckman+04}. We have shown that this would require accretion
    rates not consistent with steady-state thermally stable accretion
    of stellar wind material, and is driven by an external supply of gas.
  \item We compare the $\Mdot$ for each of our solution to the one
    calculated from the Bondi formula. We find that they agree to
    within a factor of order unity. This is because in our model the
    accretion rate is determined by the flow properties at the
    stagnation radius, $\rs$, which is within a factor of a few of the
    Bondi radius $\rb$. Thus, use of the Bondi formula is a
    good approximation.
  \end{enumerate}
